Measurement of nuclear excitation functions for proton induced reactions (E$_{\text{p}}$ = 40 -- 90 MeV) on Nb and Cu

### 2017-11-25 nb_px_paper/nb_px_paper.tex
Finished section 2.2

@@

Measurements of isomer-to-ground state ratios have been used for over 20 years to probe the spin distribution of excited nuclear states in the A$\approx$ 190 region \cite{PhysRevC.73.034613,PhysRevC.45.1171}.
In addition to the interest in the production of \ce{^{90}Mo} as an intermediate-energy dosimetry standard, this experiment offers an opportunity to study the distribution of angular momentum in compound nuclear and direct pre-equilibrium reactions via observation of a number of isomer-to-ground state ratios.
These include the \ce{^{52\text{m}}Mn} ($t_{1/2}=21.1\pm0.2$ m; J$^\pi=2^+$) to \ce{^{52\text{g}}Mn}  ($t_{1/2}=5.591\pm0.003$ d; J$^\pi=6^+$), \ce{^{58\text{m}}Co} ($t_{1/2}=9.10\pm0.09$ h; J$^\pi=5^+$) to \ce{^{58\text{g}}Co}  ($t_{1/2}=70.86\pm0.06$ d; J$^\pi=2^+$),  \ce{^{85\text{m}}Y} ($t_{1/2}=4.86\pm0.13$ h; J$^\pi=\sfrac{9}{2}^+$) to \ce{^{85\text{g}}Y}  ($t_{1/2}=2.68\pm0.05$ h; J$^\pi=\sfrac{1}{2}^-$),  \ce{^{87\text{m}}Y} ($t_{1/2}=13.37\pm0.03$ h; J$^\pi=\sfrac{9}{2}^+$) to \ce{^{87\text{g}}Y}  ($t_{1/2}=79.8\pm0.3$ h; J$^\pi=\sfrac{1}{2}^-$),  and \ce{^{89\text{m}}Nb} ($t_{1/2}=66\pm2$ m; J$^\pi=\sfrac{1}{2}^-$) to \ce{^{89\text{g}}Nb}  ($t_{1/2}=2.03\pm0.07$ h; J$^\pi=\sfrac{9}{2}^+$)  ratios \cite{Dong2015,Nesaraja2010,Singh2014,Johnson2015,Singh2013}.

@@

@@

The well-known stacked-target design was utilized for this work, in order that the (p,x) cross sections for each reaction channel could be measured at multiple energy positions in a single irradiation \cite{Graves2016}. 
For targets, a series of nominal 25 \micro m \ce{^{nat}Nb} foils (99.8\%, lot \#T23A035), 25 \micro m \ce{^{nat}Al} foils (99.999\%, lot \#M06C032), and 50 \micro m \ce{^{nat}Cu} foils (99.9999\%, lot \#N26B062) were used (all from Alfa Aesar, Ward Hill, MA, 01835, USA).
Six foils of each metal were cut down to 2.5 x 2.5 cm squares and characterized - for each foil, length and width measurements were taken at four different locations using a Mitutoyo caliper, thickness measurements were taken at four different locations using a Mitutoyo micrometer, and four mass measurements were taken using an analytical balance after cleaning the foils with isopropyl alcohol.
Using these length, width, and mass readings, the areal density (in mg/cm$^2$) for each foil was calculated, along with the propagated uncertainty in areal density.
The foils were tightly sealed into \enquote{packets} using two pieces of  3M 5413-Series Kapton polyimide film tape - each piece of tape consists of 43.2 \micro m of a silicone adhesive on 25.4 \micro m of a polyimide backing.
The sealed foils were mounted over the hollow center of a 1.575 mm-thick plastic frame, and one \ce{^{nat}Al}, one \ce{^{nat}Cu}, and \ce{^{nat}Nb} mounted foil were bundled together using baling wire for each energy position.
These foil packet bundles were lowered into the beam line by inserting them into a  water-cooled production target box.
% These six foil packet bundles were inserted into one of six  
The box, seen in \autoref{fig:target_stack}, is machined from 6061 aluminum alloy, has a thin (0.64 mm) Inconel beam entrance window, and  contains 6 \enquote{energy positions} for targets, formed by  5 slabs of 6061 aluminum alloy (previously characterized) which serve as proton energy degraders  between energy positions.
At both the front and rear of the target stack's foils, a 316 stainless steel foil is inserted to serve as a beam profile monitor - after end-of-beam (EoB), $\beta$ particles emitted from these activated stainless steel foils may be used to develop Gafchromic film, revealing the spatial profile of the beam entering and exiting the stack.
After loading all targets in the stack, the lid of the target box is sealed in place, using an inset o-ring to create a water-tight seal, and the box is lowered through a hot cell into the beamline, where it sits electrically isolated.
The specifications of the target stack design for this work is presented in \autoref{tab:stack_table}.

@@

\begin{table}[h]

@@

\begin{figure}[h]
 \centering
 \includegraphics[width=6in]{./figures/sample_gspec.pdf}
 % sample_gspec.pdf: 489x257 pixel, 72dpi, 17.25x9.07 cm, bb=0 0 489 257
 \caption{Example gamma spectra, collected from an activated Nb foil at approximately 80 MeV, to monitor radioisotope production. While the majority of observed reaction products are visible in this spectrum, it is particularly noteworthy that the \ce{^{90}Mo} decay lines which form the basis of the \ce{^{93}Nb}(p,x)\ce{^{90}Mo} monitor reaction are high in intensity and clearly isolated from surrounding peaks, greatly aiding its activity quantification.}
 \label{fig:gspec}
\end{figure}

Following  acquisition, each observed peak in the collected spectra was

identify all peaks, first observable in mass chain  = cumulative, all others are indepp

These efficiencies, along with gamma ray intensities for each transition, were used to convert the integrated counts in each gamma ray photopeak into an activity for the activated isotopes and isomeric states.

@@

\section{Decay data} \label{data}
% 
% % \begin{longtable}{|c|c|c|c|c|} 
% % \caption{My caption}
% % \label{tab:dummy}
% % % \begin{tabular}{|c|c|c|c|c|}
% %    \hline
% %  
% % \end{longtable}
Table of decay data used for observed gamma rays \cite{Wang2017,Dong2015,Dong2014,JUNDE2008787,Junde2011,Bhat1998,Nesaraja2010,BAGLIN2002,Browne2013,Zuber20151,NICHOLS2012973,Singh2007,Browne2010,Tuli2003,McCutchan2015,Singh2014,NEGRET20151,Johnson2015,McCutchan2014,Singh2013,Browne1997,Baglin2013,Baglin2012}.

@@

Plots of the cross sections measured in this work are presented here, in comparison with literature data and reaction modeling codes \cite{Koning2012,Albouy1963,PhysRev.162.1055,PhysRevC.6.1235,Grutter1982,Greenwood1984,Aleksandrov1987,levkovski1991cross,Mills1992,MICHEL1997153,Fassbender1997,Ido2002,sisterson2002selected,YashimaH2003,A2006,Ditroi2008,Ditroi2009,steyn2011excitation,Titarenko2011,Shahid2015,Garrido2016,Graves2016}.

>>>>

\newcommand{\etal}{\emph{et al.}}

@@

Measurements of isomer-to-ground state ratios have been used for over 20 years to probe the spin distribution of excited nuclear states in the A $\approx$ 190 region \cite{PhysRevC.73.034613,PhysRevC.45.1171}.
In addition to the interest in the production of \ce{^{90}Mo} as an intermediate-energy dosimetry standard, this experiment offers an opportunity to study the distribution of angular momentum in compound nuclear and direct pre-equilibrium reactions via observation of a number of isomer-to-ground state ratios.
These include the \ce{^{52\text{m}}Mn} ($t_{1/2}=21.1\pm0.2$ m; J$^\pi=2^+$) to \ce{^{52\text{g}}Mn}  ($t_{1/2}=5.591\pm0.003$ d; J$^\pi=6^+$), \ce{^{58\text{m}}Co} ($t_{1/2}=9.10\pm0.09$ h; J$^\pi=5^+$) to \ce{^{58\text{g}}Co}  ($t_{1/2}=70.86\pm0.06$ d; J$^\pi=2^+$),  \ce{^{85\text{m}}Y} ($t_{1/2}=4.86\pm0.13$ h; J$^\pi=\sfrac{9}{2}^+$) to \ce{^{85\text{g}}Y}  ($t_{1/2}=2.68\pm0.05$ h; J$^\pi=\sfrac{1}{2}^-$),  \ce{^{87\text{m}}Y} ($t_{1/2}=13.37\pm0.03$ h; J$^\pi=\sfrac{9}{2}^+$) to \ce{^{87\text{g}}Y}  ($t_{1/2}=79.8\pm0.3$ h; J$^\pi=\sfrac{1}{2}^-$),  and \ce{^{89\text{m}}Nb} ($t_{1/2}=66\pm2$ m; J$^\pi=\sfrac{1}{2}^-$) to \ce{^{89\text{g}}Nb}  ($t_{1/2}=2.03\pm0.07$ h; J$^\pi=\sfrac{9}{2}^+$)  ratios \cite{Dong2015,Nesaraja2010,Singh2014,Johnson2015,Singh2013}.

@@

While the reactions and mass region being explored in this work differ from the recent work of Graves \etal, much of the analysis described in the present work follows their established methods for the measurement of cross sections in stacked target irradiations   \cite{Graves2016}.

@@

The well-known stacked-target design was utilized for this work, in order that the (p,x) cross sections for each reaction channel could be measured at multiple energy positions in a single irradiation. 
For targets, a series of nominal 25 \micro m \ce{^{nat}Nb} foils (99.8\%, lot \#T23A035), 25 \micro m \ce{^{nat}Al} foils (99.999\%, lot \#M06C032), and 50 \micro m \ce{^{nat}Cu} foils (99.9999\%, lot \#N26B062) were used (all from Alfa Aesar, Ward Hill, MA, 01835, USA).
Six foils of each metal were cut down to 2.5 x 2.5 cm squares and characterized - for each foil, length and width measurements were taken at four different locations using a Mitutoyo caliper, thickness measurements were taken at four different locations using a Mitutoyo micrometer, and four mass measurements were taken using an analytical balance after cleaning the foils with isopropyl alcohol.
Using these length, width, and mass readings, the areal density (in mg/cm$^2$) for each foil was calculated, along with the propagated uncertainty in areal density.
The foils were tightly sealed into \enquote{packets} using two pieces of  3M 5413-Series Kapton polyimide film tape - each piece of tape consists of 43.2 \micro m of a silicone adhesive on 25.4 \micro m of a polyimide backing.
The sealed foils were mounted over the hollow center of a 1.575 mm-thick plastic frame, and one \ce{^{nat}Al}, one \ce{^{nat}Cu}, and \ce{^{nat}Nb} mounted foil were bundled together using baling wire for each energy position.
These foil packet bundles were lowered into the beam line by inserting them into a  water-cooled production target box.
% These six foil packet bundles were inserted into one of six  
The box, seen in \autoref{fig:target_stack}, is machined from 6061 aluminum alloy, has a thin (0.64 mm) Inconel beam entrance window, and  contains 6 \enquote{energy positions} for targets, formed by  5 slabs of 6061 aluminum alloy (previously characterized) which serve as proton energy degraders  between energy positions.
At both the front and rear of the target stack's foils, a 316 stainless steel foil is inserted to serve as a beam profile monitor - after end-of-beam (EoB), $\beta$ particles emitted from these activated stainless steel foils may be used to develop Gafchromic film, revealing the spatial profile of the beam entering and exiting the stack.
After loading all targets in the stack, the lid of the target box is sealed in place, using an inset o-ring to create a water-tight seal, and the box is lowered through a hot cell into the beamline, where it sits electrically isolated.
The specifications of the target stack design for this work is presented in \autoref{tab:stack_table}.

@@

\begin{table}[ht]

@@

\begin{figure}[ht]
 \centering
 \includegraphics[width=6in]{./figures/sample_gspec.pdf}
 % sample_gspec.pdf: 489x257 pixel, 72dpi, 17.25x9.07 cm, bb=0 0 489 257
 \caption{Example gamma spectra, collected from an activated Nb foil at approximately 80 MeV, to monitor radioisotope production. While the majority of observed reaction products are visible in this spectrum, it is particularly noteworthy that the \ce{^{90}Mo} decay lines which form the basis of the \ce{^{93}Nb}(p,x)\ce{^{90}Mo} monitor reaction are high in intensity and clearly isolated from surrounding peaks, greatly aiding its activity quantification.}
 \label{fig:gspec}
\end{figure}

Following  acquisition, the decaying product nuclei corresponding to each observed peak in the collected spectra were identified.
The calibrated detector efficiencies, along with gamma ray intensities for each transition, were used to convert the net  counts in each fitted gamma ray photopeak into an activity for the activated isotopes and isomeric states.
For reference, the nuclear decay data used in this work is tabulated in \autoref{tab:nudat_table_monitors} and \autoref{tab:nudat_table_nb} of \ref{data}.
As nearly all of the product nuclei created in this work have multiple high-intensity gamma rays, each product nuclei had measurements of its present activity at multiple points in time (up to 2 weeks after EoB), as well as multiple independent activity measurements at each time point, based on each of its observed gamma rays.
The measured activity for each gamma ray possesses a total propagated uncertainty which is the quadrature sum of the uncertainty in  fitted peak areas, uncertainty in detector efficiency calibration, and uncertainty in the gamma ray branching ratio data.

\comment{Add 1st and 2nd order bateman stuff here, maybe cite Cetnar2006 also }

Since many of the reaction products populated by energetic protons are more than one decay off of stability, many of these are produced not only  directly by reactions, but also indirectly produced by decay down a mass chain.
To this end, it is useful to differentiate between the types of cross sections reported in this work. 
For the first observable product nuclei in a mass chain, its (p,x) cross section will be reported as a cumulative cross section ($\sigma_c$), which is the sum of direct production of that nucleus, as well as decay of its decay precursors and any other independent cross sections leading to that nucleus. 
In addition, cumulative cross sections will be reported whenever it is impossible to use decay spectroscopy to distinguish direct production of a nucleus from decay feeding.
For all remaining observed reaction products in the mass chain, independent cross sections ($\sigma_i$) will be reported, leaving solely direct nucleus production by subtracting out the feeding of decay precursors.

To convert these activities into cross sections, corrections must be made for the decay of the various reaction products in  between EoB and the spectrum acquisition, in order to calculate $A_0$, the initial activity at EoB.
The use of  multiple gamma rays at multiple points after EoB to calculate initial activities  for each observed product nucleus allows for a more accurate  determination of $A_0$ than simply basing its calculation off of a single gamma ray observation.
For the case of cumulative cross sections, EoB activities were quantified by fitting the activities observed at multiple time points to the well-known radioactive decay law.
% \begin{equation}
% A\pp{t} = A_0 e^{-\lambda t}
% \end{equation}
Nonlinear regression was used for this fitting process, minimizing on $\chi^2$ / degree of freedom, so that not only would the uncertainty-weighted EoB activities be fitted, but that a 1-$\sigma$ confidence interval in $A_0$ would be reported as as well.
As with the gamma ray intensities, all lifetimes used in this work are tabulated in \autoref{tab:nudat_table_monitors} and \autoref{tab:nudat_table_nb} of \ref{data}.
In the case of independent cross sections, a similar process was followed, instead regressing to the solutions to the Bateman equation:
\begin{equation}
A_n\pp{t} = \lambda_n \sum_{i=1}^n \left[  N_i\pp{0} \times \pp{\prod_{j=i}^{n-1}\lambda_j} \times \pp{\sum_{j=i}^n \dfrac{e^{-\lambda_j t}}{\prod_{i\neq j}^n \pp{\lambda_i - \lambda_j}}  }   \right]
\end{equation}
While higher-order terms were added if needed, typically for an isomeric state in a particular mass chain,  the second-order expansion ($n=2$) was often sufficient to quantify EoB activities in a mass chain, simplifying to:
\begin{equation}
A_2\pp{t} = \dfrac{A_{1,0}\lambda_2}{\lambda_1 - \lambda_2} \pp{e^{-\lambda_2 t} - e^{-\lambda_1 t}} + A_{2,0} e^{-\lambda_2 t}
\end{equation}
In these cases, the previously-quantified EoB activities from decay precursors ($A_{1,0}$, etc) would be substituted in, that the feeding contributions from decay could be separated and an independent cross section reported.
After quantifying the cumulative EoB activities at the top of a mass chain, and all subsequent independent EoB activities, these would be later used to report the various cross sections for all observed reaction products and isomeric states.

@@

\section{Decay data} \label{data}
% 
% % \begin{longtable}{|c|c|c|c|c|} 
% % \caption{My caption}
% % \label{tab:dummy}
% % % \begin{tabular}{|c|c|c|c|c|}
% %    \hline
% %  
% % \end{longtable}
Table of decay data  for observed gamma rays. 
The  values for lifetime and gamma ray branching ratios  listed here were used in all calculations of measured cross sections \cite{Wang2017,Dong2015,Dong2014,JUNDE2008787,Junde2011,Bhat1998,Nesaraja2010,BAGLIN2002,Browne2013,Zuber20151,NICHOLS2012973,Singh2007,Browne2010,Tuli2003,McCutchan2015,Singh2014,NEGRET20151,Johnson2015,McCutchan2014,Singh2013,Browne1997,Baglin2013,Baglin2012}.

@@

Figures of the cross sections measured in this work are presented here, in comparison with literature data \cite{Albouy1963,PhysRev.162.1055,PhysRevC.6.1235,Grutter1982,Greenwood1984,Aleksandrov1987,levkovski1991cross,Mills1992,MICHEL1997153,Fassbender1997,Ido2002,sisterson2002selected,YashimaH2003,A2006,Ditroi2008,Ditroi2009,steyn2011excitation,Titarenko2011,Shahid2015,Garrido2016,Graves2016}, the TENDL-2015 data library \textred{(cite me!)}, and the reaction modeling codes EMPIRE-3.2.3 and TALYS-1.8 \cite{Koning2012}.
